Equations & R Code

### 2017-03-11 manuscript/coexistence-stability_ms.Rmd
adds four mus results for relnonlin SI and updates SI text
revision writing

The gap between theoretical expectations and empirical results of diversity-\new{ecosystem variability} relationships might reflect the divergence of theory developed to explain species coexistence and theory developed to explain diversity and ecosystem variability.
One reason these two disciplines have diverged is because they have focused on different questions.
Diversity-\new{ecosystem variability} studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how species richness responds to different levels of environmental variability [@Chesson1981a].

@@

\new{To bound the problem and to isolate the effect of fluctation-dependent coexistence on the diversity--ecosystem variability relationship we constrain our analysis to four species communities and do not consider communities where coexistence is maintained by a combination of mechanisms.
We limit our analysis to four species communities to keep the models tractable while still allowing enough of a species richness gradient for trends to emerge.
It is notoriously difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning} [@Yuan2015], \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities as well (see SI section SI.2).}

@@

\new{At the end of the growing season ($t=T$), a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($1-\eta_i$) to the next growing season, giving the following equation:}
\vspace{-3em}
\begin{align}
D_{i,\tau + 1} = \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right](1 - \eta_i)
\end{align}
\vspace{-3em}
\noindent \new{where $\tau(T)$ denotes the end of growing season $\tau$.
We assume remaining live biomass $\left(N_{\tau(T)}(1-\alpha)\right)$ dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).}

@@

### Numerical simulations
To explore how fluctuation-dependent coexistence can affect the diversity-variability relationship, we simulated the model with four species under two scenarios for each coexistence mechanism.
First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
To do this, we simulated communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
Second, we chose parameter values that allowed coexistence of all four species and then performed species removals, akin to a biodiversity--ecosystem function experiment.
The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.

@@

Under relative nonlinearity, species' resource response curves (Fig. S1-1) reflect traits that determine the temporal variability of each species' population growth.
"Stable" species achieve maximum resource uptake at low resource levels, but their maximum uptake rates are modest. 
For these species, population responses to resource fluctuations are buffered.
"Unstable" species have very high maximum uptake rates, which they only achieve when resource availability is high, leading to large population fluctuations.
The difference in the intrinsic stability of these two kinds of species makes our simulations sensitive to initial conditions.
Therefore, we ran two sets of simulations for relative nonlinearity: beginning with either stable or unstable species as a reference point
For example, if species A is the most stable species and species D is the least stable, we ran simulations where A then B then C then D were allowed to invade.
We then ran simulations with that order reversed.

All simulations were run for 5,000 growing seasons of 100 days each. 
We averaged biomass over the growing season, and those yearly values were used to calculate total community biomass in each year.
After discarding an initial 500 seasons to reduce transient effects on our results, we calculated the coefficient of variation (*CV*) of summed species biomass through time, which represents ecosystem variability, the inverse of ecosystem stability.
We calculated species richness as the number of species whose average biomass was greater than 1 over the course of the simulation.
Parameter values for specific results are given in figure captions.
Within-season dynamics were solved given initial conditions using the package ``deSolve`` [@Soetaert2010] in ```R``` [@Team2013].
```R``` code for our model function is in the Supporting Information section SI.1.
All model code has been deposited on Figshare (*link*) and is available on GitHub at http://github.com/atredennick/Coexistence-Stability.

RESULTS
-------
When we allowed the variance of the environment to determine which of four initial species coexisted, similar to a study across a natural diversity gradient, we found a positive relationship between richness and \new{ecosystem variability}, defined as the temporal *CV* of total community biomass (Fig. 2A,C).
This was true for the storage effect, where coexistence is maintained by fluctuating dormant-to-live transition rates ($\gamma$), and for relative nonlinearity, where coexistence is maintained by annual resource pulses.
The relationship is driven by the fact that increasing environmental variability increases the strength of both coexistence mechanisms (Fig. S1-2).
More variable conditions promoted species richness, creating a positive relationship between diversity and \new{ecosystem} variability.

When we performed species removals but held environmental variability at a level that allows coexistence of all four species, similar to a biodiversity--ecosystem functioning experiment, we found a negative diversity-\new{ecosystem variability} relationship (Fig. 2B,D).
Scatter around the relationship was small for the storage effect because all species have similar temporal variances.
Regardless of species identity, the presence of more species always stabilized ecosystem functioning through portfolio effects.
In contrast, scatter around the relationship was larger for relative nonlinearity (Fig. 2D) because species with different resource uptake curves had different population variances.
Depending on which species were present, two-species communities were sometimes less variable than three-species communities.

To understand how much species additions can stabilize ecosystem functioning as environmental variability increases, we simulated our model over a range of environmental variance and species pool sizes.
Realized species richness increased with environmental variability and, in some cases, increases in richness completely offset the effect of increasing environmental variability on ecosystem variability.
More species rich communities were less variable on average and increased in ecosystem *CV* at a slower rate than communities with fewer species (e.g., lower slopes in log-log space; Fig. S1-3).

@@

In communities where species coexist via relative nonlinearity, the extent to which species additions buffer ecosystem stability against increases in environmental variability depends on the species traits of immigrating species and the order in which they enter the community.
When additional species, which immigrate from the regional pool, are less intrinsically stable than the resident species, ecosystem variability increases at a constant rate even as species are added (Fig. 4A).
If more stable species colonize, species additions buffer the ecosystem from increasing environmental variability (Fig. 4B).
When only the most unstable (resource acquisitive) species is present, ecosystem variability initially declines as resource variability increases (Fig. 4B) because for this species' monoculture biomass is very low at low average resource levels for this species, inflating the *CV*.

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=2.5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure3.png}
  \caption{The effect of environmental variability on ecosystem variability when species coexist via relative nonlinearity. (A) The species pool increases from one to four species, with the fourth species being most unstable (e.g., resource conservative to resource acquisitive). Increasing environmental variability (the SD of annual resource availability) allows for greater species richness, but species additions do not modulate the effect of environmental variability on ecosystem variability. (B) The species pool increases from one to four species, with the fourth species being most stable (e.g., resource acquisitive to resrouce conservative). In this case, increasing environmental variability allows for greater realized species richness and can temper the effect of environmental variability. In (B), there is limited parameter space under which only three species coexist, which is why a three species community is missing. Parameter values are as in Figure 1C.}
\end{figure}

>>>>

The gap between theoretical expectations and empirical results of diversity-\new{ecosystem variability} relationships might reflect the divergence of theory developed to explain species coexistence and theory developed to explain diversity and ecosystem variability.
\new{In his thorough review of the topic, Loreau (2010) cautions that "one of the pieces of the stability jigsaw that is still missing here is the interconnection between community stability and the maintenance of species diversity due to temporal environmental variability."}
One reason these two disciplines have diverged is because they have focused on different questions.
Diversity-\new{ecosystem variability} studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how species richness responds to different levels of environmental variability [@Chesson1981a].

@@

\new{To bound the problem and to isolate the effect of fluctation-dependent coexistence on the diversity--ecosystem variability relationship we constrain our analysis to four species communities and do not consider communities where coexistence is maintained by a combination of mechanisms.
We limit our analysis to four species communities to keep the models tractable while still allowing enough of a species richness gradient for trends to emerge.
It is notoriously difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning} [@Yuan2015], \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
So, for consistency we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities too (see SI section SI.2).}

@@

\new{At the end of the growing season ($t=T$), a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($1-\eta_i$) to the next growing season, giving the following equation:}
\vspace{-3em}
\begin{align}
D_{i,\tau + 1} = \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right](1 - \eta_i)
\end{align}
\vspace{-3em}
\noindent \new{where $\tau(T)$ denotes the end of growing season $\tau$.
We assume remaining live biomass $\left(N_{\tau(T)}(1-\alpha)\right)$ dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).
We do not include extinction thresholds, or any other form of demographic stochasticity, under the assumption that we are working with abundant species with generous seed dispersal.}

@@

### Numerical simulations
To explore how fluctuation-dependent coexistence can affect the diversity-variability relationship, we simulated the model with four species under two scenarios for each coexistence mechanism.
First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
\new{To do this, we simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 values of $\sigma_R$ in the range [0, 1.4]) or environmental cue variability for the storage effect (100 values of $\sigma_E^2$ in the range [0, 3]).}
Second, we chose parameter values that allowed coexistence of all four species and then performed species removals \new{at a single level of environmental variability}, akin to a biodiversity--ecosystem function experiment.
The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.
\new{We refer to the former as a "regional" relationship, and the latter as a "local" relationship.
But note that we do not attribute any particular area size to "region", it is simply any area over which a gradient of environmental variability can emerge.}

@@

Under relative nonlinearity, species' resource response curves (Fig. S1-1) reflect traits that determine the temporal variability of each species' population growth.
"Stable" species achieve maximum resource uptake at low resource levels, but their maximum uptake rates are modest. 
For these species, population responses to resource fluctuations are buffered.
"Unstable" species have very high maximum uptake rates, which they only achieve when resource availability is high, leading to large population fluctuations.
The difference in the intrinsic stability of these two kinds of species makes our simulations sensitive to initial conditions.
Therefore, we ran two sets of simulations for relative nonlinearity: beginning with either stable or unstable species as a reference point
For example, if species A is the most stable species and species D is the least stable, we ran simulations where A then B then C then D were \new{added to the initial pool of species}.
We then ran simulations with that order reversed.

All simulations were run for 5,000 growing seasons of 100 days each. 
We averaged biomass over the growing season, and yearly values of \new{live-state biomass} were used to calculate total community biomass in each year.
After discarding an initial 500 seasons to reduce transient effects on our results, we calculated the coefficient of variation (*CV*) of summed species biomass through time, which represents ecosystem variability, the inverse of ecosystem stability.
We calculated \new{realized} species richness as the number of species whose average biomass was greater than 1 over the course of the simulation.
\new{In some cases, realized species richness is less than number of species initialized for a simulation because of competitive exclusion.}

\new{For parameters that we did not vary, we chose values that would allow coexistence of all four species at some point along the environmental variability gradients we simulated.
Our focus is specifically on communities where fluctuation-dependent coexistence is operating, and making parameters asymmetric among species typically reduced coexistence strength or made coexistence impossible (Supporting Information section SI.2).
Changes in the absolute values of parameters also altered the strength of coexistence, but in no case did altering parameter values change the qualitative results and conclusions we present below (see Supporting Information section SI.2).}
Parameter values for specific results are given in figure captions.

Within-season dynamics were solved given initial conditions using the package ``deSolve`` [@Soetaert2010] in ```R``` [@Team2013].
```R``` code for our model function is in the Supporting Information section SI.1.
All model code has been deposited on Figshare (*link*) and is available on GitHub at http://github.com/atredennick/Coexistence-Stability.

RESULTS
-------
When we allowed the variance of the environment to determine which of four initial species coexisted, similar to a study across a natural diversity gradient, we found a positive relationship between richness and \new{ecosystem variability}, defined as the temporal *CV* of total community biomass (Fig. 2A,C).
This was true for the storage effect, where coexistence is maintained by fluctuating dormant-to-live transition rates ($\gamma$), and for relative nonlinearity, where coexistence is maintained by annual resource pulses.
The relationship is driven by the fact that increasing environmental variability increases the strength of both coexistence mechanisms (Fig. S1-x).
More variable conditions promoted species richness, creating a positive relationship between diversity and \new{ecosystem} variability.

When we performed species removals but held environmental variability at a level that allows coexistence of all four species, similar to a biodiversity--ecosystem functioning experiment, we found a negative diversity-\new{ecosystem variability} relationship (Fig. 2B,D).
Scatter around the relationship was small for the storage effect because all species have similar temporal variances.
Regardless of species identity, the presence of more species always stabilized ecosystem functioning through portfolio effects.
In contrast, scatter around the relationship was larger for relative nonlinearity (Fig. 2D) because species with different resource uptake curves had different population variances.
Depending on which species were present, two-species communities were sometimes less variable than three-species communities.
\new{Furthermore, the slope of the relative nonlinearity diversity--ecosystem variability relationshp in Fig. 2D is sensitive to species' traits: the difference among species' resource uptake determines the spread of single-species communities along the y-axis.
This means that the relationship can become flat as species become more similar.}

To understand how much species additions can stabilize ecosystem functioning as environmental variability increases, we simulated our model over a range of environmental variance and species pool sizes.
Realized species richness increased with environmental variability and, in some cases, increases in richness completely offset the effect of increasing environmental variability on ecosystem variability (Fig. 3 and 4).
More species rich communities were less variable on average and increased in ecosystem *CV* at a slower rate than communities with fewer species (e.g., lower slopes in log-log space; Fig. S1-3). 
\new{The buffering effect of species richness under the storage effect is also evident in Fig. 2A because the relationship between species richness and ecosystem \emph{CV} begins to saturate.
In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. S1-x).}

@@

In communities where species coexist via relative nonlinearity, the extent to which species additions buffer ecosystem stability against increases in environmental variability depends on the species traits of immigrating species and the order in which they enter the community.
When additional species, which immigrate from the regional pool, are less intrinsically stable than the resident species, ecosystem variability increases at a constant rate even as species are added (Fig. 4A).
If more stable species colonize, species additions buffer the ecosystem from increasing environmental variability (Fig. 4B).

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=2.5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure3.png}
  \caption{The effect of environmental variability on ecosystem variability when species coexist via relative nonlinearity. (A) The species pool increases from one to four species, with the fourth species being most unstable (e.g., resource conservative to resource acquisitive). Increasing environmental variability (the SD of annual resource availability) allows for greater species richness, but species additions do not modulate the effect of environmental variability on ecosystem variability. (B) The species pool increases from one to four species, with the fourth species being most stable (e.g., resource acquisitive to resrouce conservative). In this case, increasing environmental variability allows for greater realized species richness and can temper the effect of environmental variability. In (B), there is limited parameter space under which only three species coexist, which is why a three species community is missing. Parameter values are as in Figure 2C.}
\end{figure}

### 2017-03-11 manuscript/coexistence-stability_ms.tex
adds CHPC acknowledgement
adds four mus results for relnonlin SI and updates SI text
revision writing

The gap between theoretical expectations and empirical results of
diversity-\new{ecosystem variability} relationships might reflect the
divergence of theory developed to explain species coexistence and theory
developed to explain diversity and ecosystem variability. One reason
these two disciplines have diverged is because they have focused on
different questions. Diversity-\new{ecosystem variability} studies
typically ask how ecosystem variability responds to different levels of
species richness at a given level of environmental variability (reviewed
in Kinzig et al. 2001; Loreau 2010), whereas coexistence studies ask how
species richness responds to different levels of environmental
variability (Chesson \& Warner 1981).

@@

\new{To bound the problem and to isolate the effect of fluctation-dependent coexistence on the diversity--ecosystem variability relationship we constrain our analysis to four species communities and do not consider communities where coexistence is maintained by a combination of mechanisms.
We limit our analysis to four species communities to keep the models tractable while still allowing enough of a species richness gradient for trends to emerge.
It is notoriously difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning}
(Yuan \& Chesson 2015),
\new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities as well (see SI section SI.2).}

@@

\noindent \new{where $\tau(T)$ denotes the end of growing season $\tau$.
We assume remaining live biomass $\left(N_{\tau(T)}(1-\alpha)\right)$ dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).}

@@

To explore how fluctuation-dependent coexistence can affect the
diversity-variability relationship, we simulated the model with four
species under two scenarios for each coexistence mechanism. First, we
allowed the variance of the environment to determine how many species
can coexist, akin to a community assembly experiment with a species pool
of four species. To do this, we simulated communities with all species
initially present across a gradient of annual resource variability (for
relative nonlinearity) or environmental cue variability (for the storage
effect). Second, we chose parameter values that allowed coexistence of
all four species and then performed species removals, akin to a
biodiversity--ecosystem function experiment. The two simulation
experiments correspond to (i) sampling ecosystem function across a
natural gradient of species richness and (ii) sampling ecosystem
function across diversity treatments within a site.

@@

Under relative nonlinearity, species' resource response curves (Fig.
S1-1) reflect traits that determine the temporal variability of each
species' population growth. ``Stable'' species achieve maximum resource
uptake at low resource levels, but their maximum uptake rates are
modest. For these species, population responses to resource fluctuations
are buffered. ``Unstable'' species have very high maximum uptake rates,
which they only achieve when resource availability is high, leading to
large population fluctuations. The difference in the intrinsic stability
of these two kinds of species makes our simulations sensitive to initial
conditions. Therefore, we ran two sets of simulations for relative
nonlinearity: beginning with either stable or unstable species as a
reference point For example, if species A is the most stable species and
species D is the least stable, we ran simulations where A then B then C
then D were allowed to invade. We then ran simulations with that order
reversed.

All simulations were run for 5,000 growing seasons of 100 days each. We
averaged biomass over the growing season, and those yearly values were
used to calculate total community biomass in each year. After discarding
an initial 500 seasons to reduce transient effects on our results, we
calculated the coefficient of variation (\emph{CV}) of summed species
biomass through time, which represents ecosystem variability, the
inverse of ecosystem stability. We calculated species richness as the
number of species whose average biomass was greater than 1 over the
course of the simulation. Parameter values for specific results are
given in figure captions. Within-season dynamics were solved given
initial conditions using the package \texttt{deSolve} (Soetaert et al.
2010) in \texttt{R} (Team 2013). \texttt{R} code for our model function
is in the Supporting Information section SI.1. All model code has been
deposited on Figshare (\emph{link}) and is available on GitHub at
\url{http://github.com/atredennick/Coexistence-Stability}.

@@

When we allowed the variance of the environment to determine which of
four initial species coexisted, similar to a study across a natural
diversity gradient, we found a positive relationship between richness
and \new{ecosystem variability}, defined as the temporal \emph{CV} of
total community biomass (Fig. 2A,C). This was true for the storage
effect, where coexistence is maintained by fluctuating dormant-to-live
transition rates (\(\gamma\)), and for relative nonlinearity, where
coexistence is maintained by annual resource pulses. The relationship is
driven by the fact that increasing environmental variability increases
the strength of both coexistence mechanisms (Fig. S1-2). More variable
conditions promoted species richness, creating a positive relationship
between diversity and \new{ecosystem} variability.

When we performed species removals but held environmental variability at
a level that allows coexistence of all four species, similar to a
biodiversity--ecosystem functioning experiment, we found a negative
diversity-\new{ecosystem variability} relationship (Fig. 2B,D). Scatter
around the relationship was small for the storage effect because all
species have similar temporal variances. Regardless of species identity,
the presence of more species always stabilized ecosystem functioning
through portfolio effects. In contrast, scatter around the relationship
was larger for relative nonlinearity (Fig. 2D) because species with
different resource uptake curves had different population variances.
Depending on which species were present, two-species communities were
sometimes less variable than three-species communities.

To understand how much species additions can stabilize ecosystem
functioning as environmental variability increases, we simulated our
model over a range of environmental variance and species pool sizes.
Realized species richness increased with environmental variability and,
in some cases, increases in richness completely offset the effect of
increasing environmental variability on ecosystem variability. More
species rich communities were less variable on average and increased in
ecosystem \emph{CV} at a slower rate than communities with fewer species
(e.g., lower slopes in log-log space; Fig. S1-3).

@@

In communities where species coexist via relative nonlinearity, the
extent to which species additions buffer ecosystem stability against
increases in environmental variability depends on the species traits of
immigrating species and the order in which they enter the community.
When additional species, which immigrate from the regional pool, are
less intrinsically stable than the resident species, ecosystem
variability increases at a constant rate even as species are added (Fig.
4A). If more stable species colonize, species additions buffer the
ecosystem from increasing environmental variability (Fig. 4B). When only
the most unstable (resource acquisitive) species is present, ecosystem
variability initially declines as resource variability increases (Fig.
4B) because for this species' monoculture biomass is very low at low
average resource levels for this species, inflating the \emph{CV}.

@@

The National Science Foundation provided funding for this work through a
Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a
CAREER award to PBA (DEB-1054040).
\new{We thank four anonymous reviewers for providing detailed and thoughtful comments that greatly improved the paper.}

>>>>

The gap between theoretical expectations and empirical results of
diversity-\new{ecosystem variability} relationships might reflect the
divergence of theory developed to explain species coexistence and theory
developed to explain diversity and ecosystem variability.
\new{In his thorough review of the topic, Loreau (2010) cautions that "one of the pieces of the stability jigsaw that is still missing here is the interconnection between community stability and the maintenance of species diversity due to temporal environmental variability."}
One reason these two disciplines have diverged is because they have
focused on different questions. Diversity-\new{ecosystem variability}
studies typically ask how ecosystem variability responds to different
levels of species richness at a given level of environmental variability
(reviewed in Kinzig et al. 2001; Loreau 2010), whereas coexistence
studies ask how species richness responds to different levels of
environmental variability (Chesson \& Warner 1981).

@@

\new{To bound the problem and to isolate the effect of fluctation-dependent coexistence on the diversity--ecosystem variability relationship we constrain our analysis to four species communities and do not consider communities where coexistence is maintained by a combination of mechanisms.
We limit our analysis to four species communities to keep the models tractable while still allowing enough of a species richness gradient for trends to emerge.
It is notoriously difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning}
(Yuan \& Chesson 2015),
\new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
So, for consistency we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities too (see SI section SI.2).}

@@

\noindent \new{where $\tau(T)$ denotes the end of growing season $\tau$.
We assume remaining live biomass $\left(N_{\tau(T)}(1-\alpha)\right)$ dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).
We do not include extinction thresholds, or any other form of demographic stochasticity, under the assumption that we are working with abundant species with generous seed dispersal.}

@@

To explore how fluctuation-dependent coexistence can affect the
diversity-variability relationship, we simulated the model with four
species under two scenarios for each coexistence mechanism. First, we
allowed the variance of the environment to determine how many species
can coexist, akin to a community assembly experiment with a species pool
of four species.
\new{To do this, we simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 values of $\sigma_R$ in the range [0, 1.4]) or environmental cue variability for the storage effect (100 values of $\sigma_E^2$ in the range [0, 3]).}
Second, we chose parameter values that allowed coexistence of all four
species and then performed species removals
\new{at a single level of environmental variability}, akin to a
biodiversity--ecosystem function experiment. The two simulation
experiments correspond to (i) sampling ecosystem function across a
natural gradient of species richness and (ii) sampling ecosystem
function across diversity treatments within a site.
\new{We refer to the former as a "regional" relationship, and the latter as a "local" relationship.
But note that we do not attribute any particular area size to "region", it is simply any area over which a gradient of environmental variability can emerge.}

@@

Under relative nonlinearity, species' resource response curves (Fig.
S1-1) reflect traits that determine the temporal variability of each
species' population growth. ``Stable'' species achieve maximum resource
uptake at low resource levels, but their maximum uptake rates are
modest. For these species, population responses to resource fluctuations
are buffered. ``Unstable'' species have very high maximum uptake rates,
which they only achieve when resource availability is high, leading to
large population fluctuations. The difference in the intrinsic stability
of these two kinds of species makes our simulations sensitive to initial
conditions. Therefore, we ran two sets of simulations for relative
nonlinearity: beginning with either stable or unstable species as a
reference point For example, if species A is the most stable species and
species D is the least stable, we ran simulations where A then B then C
then D were \new{added to the initial pool of species}. We then ran
simulations with that order reversed.

All simulations were run for 5,000 growing seasons of 100 days each. We
averaged biomass over the growing season, and yearly values of
\new{live-state biomass} were used to calculate total community biomass
in each year. After discarding an initial 500 seasons to reduce
transient effects on our results, we calculated the coefficient of
variation (\emph{CV}) of summed species biomass through time, which
represents ecosystem variability, the inverse of ecosystem stability. We
calculated \new{realized} species richness as the number of species
whose average biomass was greater than 1 over the course of the
simulation.
\new{In some cases, realized species richness is less than number of species initialized for a simulation because of competitive exclusion.}

\new{For parameters that we did not vary, we chose values that would allow coexistence of all four species at some point along the environmental variability gradients we simulated.
Our focus is specifically on communities where fluctuation-dependent coexistence is operating, and making parameters asymmetric among species typically reduced coexistence strength or made coexistence impossible (Supporting Information section SI.2).
Changes in the absolute values of parameters also altered the strength of coexistence, but in no case did altering parameter values change the qualitative results and conclusions we present below (see Supporting Information section SI.2).}
Parameter values for specific results are given in figure captions.

Within-season dynamics were solved given initial conditions using the
package \texttt{deSolve} (Soetaert et al. 2010) in \texttt{R} (Team
2013). \texttt{R} code for our model function is in the Supporting
Information section SI.1. All model code has been deposited on Figshare
(\emph{link}) and is available on GitHub at
\url{http://github.com/atredennick/Coexistence-Stability}.

@@

When we allowed the variance of the environment to determine which of
four initial species coexisted, similar to a study across a natural
diversity gradient, we found a positive relationship between richness
and \new{ecosystem variability}, defined as the temporal \emph{CV} of
total community biomass (Fig. 2A,C). This was true for the storage
effect, where coexistence is maintained by fluctuating dormant-to-live
transition rates (\(\gamma\)), and for relative nonlinearity, where
coexistence is maintained by annual resource pulses. The relationship is
driven by the fact that increasing environmental variability increases
the strength of both coexistence mechanisms (Fig. S1-x). More variable
conditions promoted species richness, creating a positive relationship
between diversity and \new{ecosystem} variability.

When we performed species removals but held environmental variability at
a level that allows coexistence of all four species, similar to a
biodiversity--ecosystem functioning experiment, we found a negative
diversity-\new{ecosystem variability} relationship (Fig. 2B,D). Scatter
around the relationship was small for the storage effect because all
species have similar temporal variances. Regardless of species identity,
the presence of more species always stabilized ecosystem functioning
through portfolio effects. In contrast, scatter around the relationship
was larger for relative nonlinearity (Fig. 2D) because species with
different resource uptake curves had different population variances.
Depending on which species were present, two-species communities were
sometimes less variable than three-species communities.
\new{Furthermore, the slope of the relative nonlinearity diversity--ecosystem variability relationshp in Fig. 2D is sensitive to species' traits: the difference among species' resource uptake determines the spread of single-species communities along the y-axis.
This means that the relationship can become flat as species become more similar.}

To understand how much species additions can stabilize ecosystem
functioning as environmental variability increases, we simulated our
model over a range of environmental variance and species pool sizes.
Realized species richness increased with environmental variability and,
in some cases, increases in richness completely offset the effect of
increasing environmental variability on ecosystem variability (Fig. 3
and 4). More species rich communities were less variable on average and
increased in ecosystem \emph{CV} at a slower rate than communities with
fewer species (e.g., lower slopes in log-log space; Fig. S1-3).
\new{The buffering effect of species richness under the storage effect is also evident in Fig. 2A because the relationship between species richness and ecosystem \emph{CV} begins to saturate.
In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. S1-x).}

@@

In communities where species coexist via relative nonlinearity, the
extent to which species additions buffer ecosystem stability against
increases in environmental variability depends on the species traits of
immigrating species and the order in which they enter the community.
When additional species, which immigrate from the regional pool, are
less intrinsically stable than the resident species, ecosystem
variability increases at a constant rate even as species are added (Fig.
4A). If more stable species colonize, species additions buffer the
ecosystem from increasing environmental variability (Fig. 4B).

@@

The National Science Foundation provided funding for this work through a
Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a
CAREER award to PBA (DEB-1054040).
\new{The support and resources from the Center for High Performance Computing at the University of Utah are gratefully acknowledged.}
\new{We thank four anonymous reviewers for providing detailed and thoughtful comments that greatly improved the paper.}
